\documentclass{article}
% Packages used
\input{latex-head.tex}
\def\TimesFont{} 
\graphicspath{{gfx/}}

\newcommand{\beginsupplement}{
  \setcounter{table}{0}  
  \renewcommand{\thetable}{S\arabic{table}} 
  \renewcommand{\theHtable}{S\arabic{table}} 
  \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}
    \renewcommand{\theHfigure}{S\arabic{figure}} 
  \setcounter{section}{0} 
  \renewcommand{\thesection}{S\arabic{section}}
} 
\font\supptitlefont=cmr12 at 16pt 
\newcommand{\attn }[1]{\textcolor{red}{ATTN: #1}}
     
\begin{document}
\title{Retrospective estimation of latent COVID-19 infections before Omicron in the \US}
\author[a,1]{Rachel Lobay}
%\author[b]{Maria Jahja}
\author[b]{Ajitesh Srivastava}
\author[c]{Ryan J.\ Tibshirani}
\author[a]{Daniel J.\ McDonald}

\affil[a]{Department of Statistics, The University of British Columbia}
%\affil[b]{Department of Statistics \& Data Science, Carnegie Mellon University}
\affil[b]{Department of Computer and Electrical Engineering, University of Southern California}
\affil[c]{Department of Statistics, The University of California, Berkeley}

\footnotetext[1]{To whom correspondence should be
  addressed. E-mail: \email{rachel.lobay@stat.ubc.ca}} 



\date{Version: \today}
\maketitle

\begin{abstract}
The true timing and magnitude of COVID-19 infections are of interest to both the
public and to public health, but these are challenging to pin down for a variety
of reasons such as the underdetection of asymptomatic cases and delays in
reporting. Accurate estimates of latent COVID-19 infections can improve our
understanding of the size and scope of the pandemic and provide more meaningful
and timely quantification of disease patterns and burden. In this work, we
describe methodology to estimate daily incident \emph{infections} for each \US
state. To this end, we first deconvolve reported COVID-19 cases to their date of
infection onset using delay distributions estimated from the CDC line list. We
then use a serology-driven model to scale these deconvolved cases to account for
the unreported infections. The resulting daily infection estimates
incorporate variant-specific incubation periods, reinfections,
and waning antigenic immunity. This analysis also produces estimates for other
important quantities such as the number of deconvolved cases specific to each
variant and the infection-case-report ratio (also referred to as the case
ascertainment rate). Finally, we discuss some implications of our results: a
disease burden that appears earlier and more extensively than previously
quantified and differential infection-hospitalization ratio estimates. Our
findings help to better understand the progression of the pandemic in the \US
prior to the onset of Omicron and its descendants. 

\end{abstract}

\input{doc/intro.tex}

\input{doc/results.tex}

\input{doc/discussion.tex}

\input{doc/methods.tex}


\section*{Data availability}

The Nationwide Commercial Laboratory Seroprevalence Survey 
and 2020-2021 Nationwide Blood Donor Seroprevalence Survey 
data are both freely available at
\href{https://data.cdc.gov/Laboratory-Surveillance/Nationwide-Commercial-Laboratory-Seroprevalence-Su/d2tw-32xv}{https://data.cdc.gov/Laboratory-Surveillance/Nationwide-Commercial-Laboratory-Seroprevalence-Su/d2tw-32xv} 
and 
\href{https://data.cdc.gov/Laboratory-Surveillance/2020-2021-Nationwide-Blood-Donor-Seroprevalence-Su/mtc3-kq6r}{https://data.cdc.gov/Laboratory-Surveillance/2020-2021-Nationwide-Blood-Donor-Seroprevalence-Su/mtc3-kq6r}. 
The COVID-19 Case Surveillance Restricted Access Detailed Data 
requires a Restriction Information and Data Use Restrictions 
Agreement. More information on the dataset and the agreement 
can be found \href{https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Restricted-Access-Detai/mbd7-r32t/about\_data}{https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Restricted-Access-Detai/mbd7-r32t/about\_data}. 
The JHU reported case and hospital admission data are 
publicly accessible through the COVIDcast API, 
\href{https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html}{https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html}. 
The data used to construct the incubation periods was obtained from \newline
\href{https://elifesciences.org/articles/57149}{https://elifesciences.org/articles/57149}, 
\href{https://pubmed.ncbi.nlm.nih.gov/35627870/}{https://pubmed.ncbi.nlm.nih.gov/35627870/}, \newline
\href{https://pubmed.ncbi.nlm.nih.gov/34849500/}{https://pubmed.ncbi.nlm.nih.gov/34849500/}, 
and \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8834809/}{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8834809/}. 
The data used to estimate the variant proportions in circulation is freely 
available at \newline \href{https://github.com/hodcroftlab/covariants}{https://github.com/hodcroftlab/covariants}. 
The prior inverse ascertainment ratio estimates were extracted from Table 1 in
 \href{https://www.nature.com/articles/s41467-020-19652-6}{https://www.nature.com/articles/s41467-020-19652-6}. 
 The fraction of new infections are based on estimates from \href{https://wwwnc.cdc.gov/eid/article/28/10/22-1045\_article}{https://wwwnc.cdc.gov/eid/article/28/10/22-1045\_article}. 
Finally, the state population estimates are from a U.S. Census Bureau press release at \href{https://www.census.gov/data/tables/time-series/demo/popest/2020s-state-total.html}{https://www.census.gov/data/tables/time-series/demo/popest/2020s-state-total.html}.




\section*{Code availability}
Code for reproducing all figures and the numerical results is 
available at 
\href{https://github.com/cmu-delphi/latent-infections/}{https://github.com/cmu-delphi/latent-infections/}.

\clearpage
%\bibliographystyle{apalike}
\bibliography{bibliography.bib}
\bibliographystyle{naturemag} %% Change back to numeric references in line with
%requirements for Nature Communications articles (see pg 3 fro here:
%https://www.nature.com/documents/ncomms-formatting-instructions.pdf)



\section*{Acknowledgements}

We would like to thank members of the Delphi research group for valuable
feedback, and Change Healthcare and Optum/United Health Group for their
invaluable data partnership and collaboration. 

We gratefully acknowledge all
data contributors, i.e., the Authors and their Originating laboratories
responsible for obtaining the specimens, and their Submitting laboratories for
generating the genetic sequence and metadata and sharing via the GISAID
Initiative \citep{elbe2017data}, on which this research is based.
% Required Gisaid acknowledgement

Any opinions, findings, and
conclusions or recommendations expressed in this material are
those of the authors and do not necessarily reflect the views of
the National Science Foundation and the Centers for Disease
Control and Prevention.
% NSF and CDC acknowledgement 

DJM and RJT were supported by Centers for Disease Control and Prevention (CDC)
Grant No.\ 75D30123C15907. DJM and RL received support from the National
Sciences and Engineering Research Council of Canada and the University of
British Columbia. AS was supported by the Centers for Disease Control
and Prevention and the National Science Foundation under
Award No.\ 2223933 and 2333494. 


\section*{Author contributions}
RT and DJM conceptualized the project. RL acquired the data, performed the
statistical analysis, and wrote the initial draft of the paper. DJM, RT, and AS,
reviewed and revised the paper. All authors were involved in significant
discussion and development of the methodology.

\section*{Competing interests}

The authors declare no competing interests.

\clearpage
% Eventually make the supplement a separate document with its own title page 
\beginsupplement
\title{\supptitlefont\noindent Supplementary Information: Retrospective estimation of latent COVID-19 infections before Omicron in the \US}
\maketitle

\input{doc/supplement.tex}

\end{document}